A thesis draft

### 2018-11-20 res/sections/08-Background.tex
Start SFC explaination

The NFV technology is exploited in a framework, organized
at least by three components~\cite{etsi2013gs}, as represented in
Figure~\ref{chap:background:img:etsi_arch}:
\begin{enumerate}
  \item a software, able to run on over the NFVI, that performs the network
  function that must be deployed, that is the VNF;
  \item NFV Infrastructure (NFVI) that is composed by all physical resources
  and how these can be virtualized, that forms the environment where the VNFs
  are deployed;
  \item NFV Management And Orchestration (NFV-MANO) that covers the
  orchestration and management of both physical and logical resources, the
  functional blocks, data repositories used by these blocks, reference points
  and interfaces through which these functional blocks exchange
  information for the purpose of managing and orchestrating NFVI and VNFs.
\end{enumerate}

@@

@@

>>>>

The NFV technology is exploited in a framework, organized
at least by three components~\cite{etsi2013gs}, as represented in
Figure~\ref{chap:background:img:etsi_arch}:
\begin{enumerate}
  \item a software, able to run on over the NFVI, that performs the network
  function that must be deployed, that is the VNF;
  \item NFV Infrastructure (NFVI) that is composed by all physical resources
  and how these can be virtualized, that forms the environment where the VNFs
  are deployed;
  \item NFV MANagement and Orchestration (NFV-MANO) that covers the
  orchestration and management of both physical and logical resources, the
  functional blocks, data repositories used by these blocks, reference points
  and interfaces through which these functional blocks exchange
  information for the purpose of managing and orchestrating NFVI and VNFs.
\end{enumerate}

@@

\subsection{NFV MANagement and Orchestration}
The MANO, represented in Figure~\ref{chap:background:img:etsimano}, can be splitted into three logical functional components:
\begin{itemize}
  \item the virtual infrastructure manager (VIM);
  \item the VNF Manager (VNFM);
  \item the NFV Orchestrator (NFVO).
\end{itemize}
And into four data repositories:
\begin{itemize}
  \item Network Service (NS) catalog;
  \item VNF catalog;
  \item NFV instances;
  \item NFVI resources.
\end{itemize}

\begin{figure}
  \centering
  \includegraphics[scale=0.3]{etsimano}
  \caption{NFV MANO components}
  \label{chap:background:img:etsimano}
\end{figure}

\paragraph*{VIM}
The VIM aim is to manage and control NFVI resources. Every MANO can contain more VIM instances, each of them control NFVI resources from a single infrastructure. VIM can be developed either to administer one type of NFVI or more. VIM have to monitor the allocation of virtual resource in order to orchestrate allocation, upgrade, release, and reclamation of resources and optimize their usage, including virtual links, networks, subnets, and ports. It is also the component in charge to manage security group policies to ensure access control. In addition, VIM also has to record data on performance and faults, that can be queried. Thank to this component it is possible to dynamically allocate resource, based on demand, that is particularly using IaaS.

\paragraph*{VNFM} 
In this framework all VNFs, once created, have to register to a VNFM. This
component is accountable to manage their lifecycle, that consists of
instantiation, scaling, termination and, possibly, update/upgrade, and
communicate the state of a certain VNF to others MANO blocks: this enable the
monitoring of key performance indicators (KPI), that can help to make decision
on VNF scaling. VNFM can manage a single type of VNF instances or multiple.

\paragraph*{NFVO}
The NFVO is responsible of combining functions to create end-to-end services.
Its functionalities are to ensure there are adequate compute, storage, and
network resources available to provide a network service independently of the
VIM and the creation of end-to-end services by composing different VNFs, that
can be deployed on different point-of-presence and registered to different VNF.
It also managing the topology of these services.

\paragraph*{Data repositories} Data repositories are use to keep information on
the NFV MANO. The NS catalog is a set of predefined templates, which define
services, functions involved and their connectivity. The VNF catalog is,
instead, a set of templates describing the deployment and characteristics of
VNFs. The NFVI resources repository holds information about available/allocated
NFVI resources. The NFV instances repository keep information about functions
and services instances during their lifetimes.

@@

\section{Service Function Chaining}
The Service Function Chaining (SFC) concept was defined  in~\cite{halpern2015service} and refers to an ordered set of Service Functions (SF) and the constraint, defined after a classification process, to packets and flows. SF are the particular functions whose aim is the treatment of packets received. They can work on different levels of the network layer. These components can be implemented as virtual element or a combination of a physical network element and the software running on the top of it. There are no limitations on the number of SFs running on the same network element, so in general it is possible to have one or more functions running on the same element. SFs can be aware or not that are part of a certain chain, in the matter case they are called \emph{SFC aware} functions, in the latter \emph{SFC unaware}.

An SFC can be seen as a graph in which each node is an SF and edges are the path through which a packet have to pass through. Because of this definitions there are no limitation on possible cycles that can occur in the chain, in case that a certain SF have to be reapplied on the same packet, frame or flow. These scenarios must be adequately managed to avoid endless loops. This enable even the possibility of branching on a certain node in order to define the treatment that has to be imposed from that point of the chain. The graph representation also permit to think to \emph{unidirectional} and \emph{bidirectional} chains. An SFC is defined unidirectional if traffic can traverse it only in one direction, so if it is composed by 3 SF, named \texttt{SF0}, \texttt{SF1} and \texttt{SF2} in this order, traffic can only go from \texttt{SF0} to \texttt{SF2}. On the contrary, a bidirectional SFC allows traffic in both directions. So, having the same chain as in the previous example, traffic can pass both from \texttt{SF0} to \texttt{SF2} and from \texttt{SF2} to \texttt{SF0}.
